Analysis

### 2025-04-30 code/analysis-code/analysis.qmd
Part 6 updates

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(survey)
library(tableone)
library(mice)
library(mitools)
library(miceadds)
library(tidymodels)
library(caret)
library(gt)
library(here)
```

@@

Next, we will make the multivariate model. I will use cross validation for this. The RMSE is found to be 0.4450602.
```{r}

cross_validation_imputed_rmse <- function(imputed_data, design, predictors, k_folds = 5) {
  rmse_list <- numeric(length(imputed_data))  # Store RMSE for each imputation

  for (i in 1:length(imputed_data)) {
    data <- imputed_data[[i]]
    set.seed(123)
    folds <- sample(1:k_folds, size = nrow(data), replace = TRUE)

    fold_rmse <- numeric(k_folds)  # RMSE for each fold

@@

      # Survey design for training and testing
      train_design <- svydesign(ids = ~1, data = train_data, weights = ~SAMPWEIGHT)
      formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
      model <- svyglm(formula, design = train_design, family = quasibinomial())

      # Predictions
      predictions <- predict(model, newdata = test_data, type = "response")
      actual <- test_data$COVID

      # Compute RMSE for this fold
      fold_rmse[fold] <- sqrt(mean((predictions - actual)^2, na.rm = TRUE))
    }

    # Average RMSE over folds for this imputed dataset
    rmse_list[i] <- mean(fold_rmse)
  }

  # Average RMSE over all imputed datasets
  overall_rmse <- mean(rmse_list)

  return(list(rmse_per_imputation = rmse_list, overall_rmse = overall_rmse))
}

# Assuming you have your imputed_data and a vector of predictors
results <- cross_validation_imputed_rmse(imputed_data = data_imputed,
                                         design = NULL,  # Not used inside, can be omitted
                                         predictors = c("AGE", "SLEEP", "DEPRESSED"))

# View results
results$rmse_per_imputation  # RMSE for each imputed dataset
results$overall_rmse         # Average RMSE across imputations
```

Now, we will make a LASSO regression model. For this, I'm not going to use the multiple imputation dataset. I'm just going to pick a sinlge imputed dataset, the first one. The RMSE is 0.276.

```{r}
# Extract the first completed imputed dataset
data1 <- data_imputed[[1]]  # You can loop over 1:5 later

# Then define the survey design
survey_design <- svydesign(
  ids = ~PSU,
  strata = ~STRATA,
  weights = ~SAMPWEIGHT,
  data = data1,
  nest=TRUE
)

@@

# Choose your outcome and predictors
outcome_var <- "DEPRESSED"
predictor_vars <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")

# Loop over imputations and compute RMSE
rmse_list <- map(data_imputed, function(dat) {
  # Drop missing in predictors or outcome if any slipped through
  dat <- dat %>% drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))

@@

  # Fit weighted lasso model
  cv_fit <- cv.glmnet(
    x = X_train,
    y = y_train,
    alpha = 1,
    weights = w_train,
    family = "binomial",  # change to "binomial" for binary outcome
    nfolds = 5
  )

  # Predict on test set
  preds <- predict(cv_fit, newx = X_test, s = "lambda.min")

  # Weighted RMSE
  sqrt(weighted.mean((y_test - preds)^2, w_test))
})

# Pool RMSE across imputations
pooled_rmse <- mean(unlist(rmse_list))

# Output
cat("RMSE for each imputation:\n")
print(unlist(rmse_list))
cat("\nPooled RMSE across imputations:", round(pooled_rmse, 3), "\n")

################################################################################
The following code was used prior to development of imputed data and is no longer relevant. It may be ignored. I am keeping it in case it serves as a useful reference for the rest of the project or future endeavors.
################################################################################

We create estimates and CIs using complex survey design variables
```{r}
#list variables
varstable1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

#list categories
cat_vartab1 = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
             "SOCIAL", "COVID")

survey = svydesign(
  ids = ~PSU, 
  strata = ~STRATA, 
  weights = ~SAMPWEIGHT, 
  nest = TRUE, 
  data = data)

exercise <- svyby(~as.factor(EXERCISE), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
sleep <- svyby(~as.factor(SLEEP), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
depressed <- svyby(~as.factor(DEPRESSED), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
trouble_sleeping <- svyby(~as.factor(TROUBLE_SLEEPING), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
social <- svyby(~as.factor(SOCIAL), ~COVID, survey, svymean, vartype = "ci", na.rm = TRUE)
str(exercise)
```

Let's rearrange these exercise numbers into a table. I used ChatGPT and Claude to help with making the table. 
```{r}
# Convert the svyby object to a data frame (if not already)
exercise_df <- as.data.frame(exercise)

# Clean up column names to make them more readable
colnames(exercise_df) <- c("COVID", "EXERCISE_0", "EXERCISE_1", "CI_L_EXERCISE_0", "CI_L_EXERCISE_1", "CI_U_EXERCISE_0", "CI_U_EXERCISE_1")

# Use a different approach to reshape the data
exercise_long <- exercise_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and exercise level
  mutate(
    Type = case_when(
      grepl("^EXERCISE", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    Exercise = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, Exercise),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
exercise_with_ci <- exercise_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, Exercise, Mean, CI)

# Now create the final table format
exercise_final <- exercise_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = Exercise,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
exercise_table <- exercise_final %>%
  gt() %>%
  tab_header(
    title = "Exercise Data by COVID Status"
  ) %>%
  cols_label(
    Exercise = "Exercise Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = Exercise == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Exercise Level: 0 = No regular exercise, 1 = Meets requirements for regular exercise according to American Heart Association",
    locations = cells_column_labels(columns = Exercise)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = Exercise),
    fn = function(x) {
      ifelse(x == "0", "No Regular Exercise", "Regular Exercise")
    }
  )

# Print the table
exercise_table
```

Next, we will add the sleep variable. I used ChatGPT for this.
```{r}
# Convert the svyby object to a data frame (if not already)
sleep_df <- as.data.frame(sleep)  # Assuming 'sleep' is the dataset

# Clean up column names to make them more readable
colnames(sleep_df) <- c("COVID", "SLEEP_0", "SLEEP_1", "CI_L_SLEEP_0", "CI_L_SLEEP_1", "CI_U_SLEEP_0", "CI_U_SLEEP_1")

# Use a different approach to reshape the data
sleep_long <- sleep_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and sleep level
  mutate(
    Type = case_when(
      grepl("^SLEEP", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    Sleep = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, Sleep),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
sleep_with_ci <- sleep_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, Sleep, Mean, CI)

# Now create the final table format
sleep_final <- sleep_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = Sleep,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
sleep_table <- sleep_final %>%
  gt() %>%
  tab_header(
    title = "Sleep Data by COVID Status"
  ) %>%
  cols_label(
    Sleep = "Sleep Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = Sleep == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Sleep Level: 0 = Less than recommended sleep, 1 = Met recommendations for sleep by age groups according to NIH",
    locations = cells_column_labels(columns = Sleep)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = Sleep),
    fn = function(x) {
      ifelse(x == "0", "Not enough sleep", "Enough sleep")
    }
  )

# Print the table
sleep_table

#################################################################################

# Extract the data from the gt tables
exercise_table_df <- as.data.frame(exercise_table)
sleep_table_df <- as.data.frame(sleep_table)

# Ensure that the column names are the same across both tables
colnames(sleep_table_df) <- colnames(exercise_table_df)

# Combine the tables row by row
combined_df <- bind_rows(exercise_table_df, sleep_table_df)

# Convert the combined data frame back to a gt table
combined_table <- combined_df %>%
  gt() %>%
  tab_header(
    title = "Combined Exercise and Sleep Data by COVID Status"
  )

# Print the combined gt table
combined_table
```

Create the same tables with depressed, trouble_sleeping, and social. This was completed in BBEdit.
```{r}
#########################################################################################
# depressed variable

# Convert the svyby object to a data frame (if not already)
depressed_df <- as.data.frame(depressed)

# Clean up column names to make them more readable
colnames(depressed_df) <- c("COVID", "depressed_0", "depressed_1", "CI_L_depressed_0", "CI_L_depressed_1", "CI_U_depressed_0", "CI_U_depressed_1")

# Use a different approach to reshape the data
depressed_long <- depressed_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and depressed level
  mutate(
    Type = case_when(
      grepl("^depressed", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    depressed = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, depressed),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
depressed_with_ci <- depressed_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, depressed, Mean, CI)

# Now create the final table format
depressed_final <- depressed_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = depressed,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
depressed_table <- depressed_final %>%
  gt() %>%
  tab_header(
    title = "depressed Data by COVID Status"
  ) %>%
  cols_label(
    depressed = "depressed Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = depressed == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Depressed level: 0 = Feeling depressed less than weekly, 1 = Feeling depressed weekly or more",
    locations = cells_column_labels(columns = depressed)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = depressed),
    fn = function(x) {
      ifelse(x == "0", "Not regularly depressed", "Frequent depressed feeling")
    }
  )

# Print the table
depressed_table

#########################################################################################
# trouble_sleeping variable

# Convert the svyby object to a data frame (if not already)
trouble_sleeping_df <- as.data.frame(trouble_sleeping)

# Clean up column names to make them more readable
colnames(trouble_sleeping_df) <- c("COVID", "trouble_sleeping_0", "trouble_sleeping_1", "CI_L_trouble_sleeping_0", "CI_L_trouble_sleeping_1", "CI_U_trouble_sleeping_0", "CI_U_trouble_sleeping_1")

# Use a different approach to reshape the data
trouble_sleeping_long <- trouble_sleeping_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and trouble_sleeping level
  mutate(
    Type = case_when(
      grepl("^trouble_sleeping", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    trouble_sleeping = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, trouble_sleeping),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
trouble_sleeping_with_ci <- trouble_sleeping_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, trouble_sleeping, Mean, CI)

# Now create the final table format
trouble_sleeping_final <- trouble_sleeping_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = trouble_sleeping,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
trouble_sleeping_table <- trouble_sleeping_final %>%
  gt() %>%
  tab_header(
    title = "trouble_sleeping Data by COVID Status"
  ) %>%
  cols_label(
    trouble_sleeping = "trouble_sleeping Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = trouble_sleeping == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "trouble_sleeping level: 0 = No trouble sleeping, 1 = Reported trouble sleeping",
    locations = cells_column_labels(columns = trouble_sleeping)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = trouble_sleeping),
    fn = function(x) {
      ifelse(x == "0", "No reported trouble sleeping", "Regular trouble sleeping")
    }
  )

# Print the table
trouble_sleeping_table

#########################################################################################
# social variable

# Convert the svyby object to a data frame (if not already)
social_df <- as.data.frame(social)

# Clean up column names to make them more readable
colnames(social_df) <- c("COVID", "social_0", "social_1", "CI_L_social_0", "CI_L_social_1", "CI_U_social_0", "CI_U_social_1")

# Use a different approach to reshape the data
social_long <- social_df %>%
  # First, let's gather all columns except COVID
  pivot_longer(
    cols = -COVID,
    names_to = "variable",
    values_to = "value"
  ) %>%
  # Now separate the variable column into type and social level
  mutate(
    Type = case_when(
      grepl("^social", variable) ~ "Mean",
      grepl("^CI_L", variable) ~ "L",
      grepl("^CI_U", variable) ~ "U"
    ),
    social = case_when(
      grepl("_0$", variable) ~ "0",
      grepl("_1$", variable) ~ "1"
    )
  ) %>%
  # Now pivot wider to get the structure we want
  pivot_wider(
    id_cols = c(COVID, social),
    names_from = Type,
    values_from = value
  )

# Combine the CI values into a single column
social_with_ci <- social_long %>%
  mutate(
    CI = paste0("(", round(L, 2), ", ", round(U, 2), ")")
  ) %>%
  select(COVID, social, Mean, CI)

# Now create the final table format
social_final <- social_with_ci %>%
  # Convert COVID to character to ensure it works as a column header
  mutate(COVID = as.character(COVID)) %>%
  # Pivot to get COVID status as columns
  pivot_wider(
    id_cols = social,
    names_from = COVID,
    values_from = c("Mean", "CI"),
    names_glue = "{.value}_{COVID}"
  )

# Create the table using gt
social_table <- social_final %>%
  gt() %>%
  tab_header(
    title = "social Data by COVID Status"
  ) %>%
  cols_label(
    social = "social Level",
    Mean_0 = "Mean",
    Mean_1 = "Mean",
    CI_0 = "95% CI",
    CI_1 = "95% CI"
  ) %>%
  tab_spanner(
    label = "COVID Negative (0)",
    columns = c(Mean_0, CI_0)
  ) %>%
  tab_spanner(
    label = "COVID Positive (1)",
    columns = c(Mean_1, CI_1)
  ) %>%
  # Add formatting for the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add some style
  tab_style(
    style = cell_fill(color = "#f5f5f5"),
    locations = cells_body(rows = social == "0")
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "social Level: 0 = Regular social support, 1 = Less social support",
    locations = cells_column_labels(columns = social)
  ) %>%
  # Add meaningful row labels
  text_transform(
    locations = cells_body(columns = social),
    fn = function(x) {
      ifelse(x == "0", "Regular social support", "Less social support")
    }
  )

# Print the table
social_table

Now, we will create the combined table of exercise, sleep, depressed, trouble_sleeping, and social. ChatGPT helped with many tweaks.
```{r}
# Extract the data from the other gt tables
depressed_table_df <- as.data.frame(depressed_table)
trouble_sleeping_table_df <- as.data.frame(trouble_sleeping_table)
social_table_df <- as.data.frame(social_table)

# Ensure the column names match across all tables
colnames(depressed_table_df) <- colnames(combined_df)
colnames(trouble_sleeping_table_df) <- colnames(combined_df)
colnames(social_table_df) <- colnames(combined_df)

# Combine all the tables row by row
combined_all_df <- bind_rows(combined_df, depressed_table_df, trouble_sleeping_table_df, social_table_df)

# Add a row_id column to the data frame
combined_all_df <- combined_all_df %>%
  mutate(row_id = row_number())

# Convert the combined data frame back to a gt table
combined_all_table <- combined_all_df %>%
  gt() %>%
  tab_header(
    title = "Mental Health Factors by SARS-CoV-2 Test Status"
  ) %>%
  # Spanners for COVID Negative and COVID Positive
  tab_spanner(
    label = "Never has tested positive for Covid",
    columns = starts_with("Estimate") | starts_with("CI 95%")
  ) %>%
  tab_spanner(
    label = "Has tested positive for Covid",
    columns = starts_with("Estimate") | starts_with("CI 95%")
  ) %>%
  # Remove the word 'Exercise' from the first row
  cols_label(
    Exercise = ""  # This removes the "Exercise" label
  ) %>%
  # Format the mean values
  fmt_number(
    columns = starts_with("Mean"),
    decimals = 2
  ) %>%
  # Add a footnote
  tab_footnote(
    footnote = "Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group.",
    locations = cells_column_labels(columns = Exercise)
  ) %>%
  # Add alternating row colors using row_id
  tab_style(
    style = cell_fill(color = "#f5f5f5"),  # Gray color for odd rows
    locations = cells_body(rows = row_id %% 2 == 1)  # Apply to odd rows
  ) %>%
  tab_style(
    style = cell_fill(color = "white"),  # White color for even rows
    locations = cells_body(rows = row_id %% 2 == 0)  # Apply to even rows
  ) %>%
  # Add solid vertical line between row names and Mean_0 columns
  tab_style(
    style = cell_borders(
      sides = c("left"),
      color = "grey",
      weight = px(1)
    ),
    locations = cells_body(
      columns = starts_with("Mean_0")
    )
  ) %>%
  # Add solid vertical line between COVID Negative and COVID Positive columns
  tab_style(
    style = cell_borders(
      sides = c("left"),
      color = "grey",
      weight = px(1)
    ),
    locations = cells_body(
      columns = starts_with("Mean_1")
    )
  ) %>%
  # Hide the row_id column
  cols_hide(columns = "row_id")

# Print the combined gt table
combined_all_table

#footnote:
#"Exercise requirements determined according to American Heart Association. Sleep levels determined by NIH recommendations by age group."

# Save the table as a PNG
here()
gtsave(combined_all_table, here("results", "figures", "result_table1.png"))
```

################################################################################

Up next, we will run the singleton models. For these, our outcome variable (covid test result) will be used as a variable for each of the respective predictor vairbales that we created in the processing code. These models appear as the following:

$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Sleep} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Depressed} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{TroubleSleeping} $$
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Social} $$
Create survey design.
```{r}
#review names to use below
names(data)

#create survey object using design variables
survey_design <- svydesign(
  ids = ~PSU,
  strata = ~STRATA,
  weights = ~SAMPWEIGHT,
  data = data,
  nest = TRUE
)

pr_results <- data.frame()

variables <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", 
               "SOCIAL")

for (var in variables) {
  # Create the formula dynamically
  formula <- as.formula(paste("COVID ~", var))

  # Fit the model
  model <- svyglm(
    formula, 
    design = survey_design, 
    family = quasibinomial(link = "log")
  )

  # Extract prevalence ratios and confidence intervals
  pr <- exp(coef(model))  # Prevalence ratios
  pr_ci <- exp(confint(model))  # Confidence intervals

  # Extract p-values
  p_values <- summary(model)$coefficients[, "Pr(>|t|)"]

  # Combine results into a single table
  pr_results <- rbind(
    pr_results,
    data.frame(
      Variable = var,
      Estimate = pr[-1],  # Skip the intercept
      `2.5% CI` = pr_ci[-1, 1],
      `97.5% CI` = pr_ci[-1, 2],
      `p-value` = p_values[-1]  # Skip the intercept
    )
  )
}

# Print the results
print(pr_results)

Now, we will make that into a nice quality table using the gt package and the help of chatGPT and Claude.
```{r}
#checking
glimpse(pr_results)
str(pr_results)

#create the gt table
pr_results %>%
  mutate(
    Predictor = stringr::str_to_title(gsub("_", " ", Variable)),
    `Prevalence Ratio` = round(Estimate, 2),
    `95% CI` = paste0("(", round(X2.5..CI, 2), ", ", round(X97.5..CI, 2), ")"),
    `p-value` = format.pval(p.value, digits = 3, eps = .001)
  ) %>%
  select(Predictor, `Prevalence Ratio`, `95% CI`, `p-value`) %>%
  gt() %>%
  tab_header(
    title = md("**Positive COVID Test Result Association with Mental Health Factors**")
  ) %>%
  tab_options(
    column_labels.border.bottom.width = px(1.5),  # Bold line under column headers
    column_labels.border.bottom.color = "black"
  )

Now, we're going to complete the process of model selection, using backwards selection. We want to end up with a model we can use for prediction purposes. Our initial model looks like this: 
$$ Y_{CovidPositive} = \beta_0 + \beta_1 \cdot X_{Exercise} + \beta_2 \cdot X_{Sleep} + \beta_3 \cdot X_{Depressed} + \beta_4 \cdot X_{TroubleSleeping} + \beta_5 \cdot X_{Social} $$
Let's find the AIC value. 
```{r}

unique(data$COVID)
model_full <- svyglm(
  COVID ~ EXERCISE + SLEEP + DEPRESSED + TROUBLE_SLEEPING + SOCIAL,
  design = survey_design,
  family = quasibinomial(link = "log")
)
# Perform backward selection based on AIC
model_step <- step(model_full, direction = "backward", trace = 1)
summary(model_step)

>>>>

To start, I'll load packages. As I need packages along the way, I'll return here and add them.
```{r}
library(tidyverse)
library(survey)
library(tableone)
library(mice)
library(mitools)
library(miceadds)
library(tidymodels)
library(caret)
library(gt)
library(pROC)
library(ranger)
library(glmnet)
library(Metrics)
library(purrr)
library(here)
```

@@

Next, we will make the multivariate model. I will use cross validation for this. The pooled AUC across imputations is found to be 0.5230.
```{r}
# Define the AUC cross-validation function for imputed datasets
cross_validation_imputed_auc <- function(imputed_data, predictors, k_folds = 5) {
  auc_list <- numeric(length(imputed_data))  # Store AUC for each imputation

  for (i in 1:length(imputed_data)) {
    data <- imputed_data[[i]]

    # drop rows with missing values in COVID, predictors, or SAMPWEIGHT
    data <- data %>% drop_na(all_of(c("COVID", predictors, "SAMPWEIGHT")))

    set.seed(123)
    folds <- sample(1:k_folds, size = nrow(data), replace = TRUE)

    fold_auc <- numeric(k_folds)  # AUC for each fold

@@

      # Fit random forest model on training data
      formula <- as.formula(paste("COVID ~", paste(predictors, collapse = " + ")))
      rf_model <- ranger(
        formula = formula,
        data = train_data,
        case.weights = train_data$SAMPWEIGHT,
        probability = TRUE
      )

      # Predict probabilities for class 1 (COVID = 1)
      pred_probs <- predict(rf_model, data = test_data)$predictions[, 2]
      actual <- test_data$COVID

      # Compute AUC for this fold
      auc_result <- roc(actual, pred_probs)$auc
      fold_auc[fold] <- auc_result
    }

    # Average AUC over folds for this imputed dataset
    auc_list[i] <- mean(fold_auc)
  }

  # Average AUC over all imputed datasets
  overall_auc <- mean(auc_list)

  return(list(auc_per_imputation = auc_list, overall_auc = overall_auc))
}

# Assuming you have your imputed_data and a vector of predictors
results <- cross_validation_imputed_auc(imputed_data = data_imputed,
                                        predictors = c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL"))

# View results
results$auc_per_imputation  # AUC for each imputed dataset
results$overall_auc         # Average AUC across imputations
```

Now, we will make a LASSO regression model. For this, I'm going to use the multiple imputation dataset with cross validation with 5 folds. Instead of using complex survey design, I will weight the model by sampleweight. The pooled AUC is  0.529.

```{r}
# Extract the first completed imputed dataset
data1 <- data_imputed[[1]]

# Define the survey design
survey_design <- svydesign(
  ids = ~PSU,
  strata = ~STRATA,
  weights = ~SAMPWEIGHT,
  data = data1,
  nest=TRUE
)

@@

# Choose outcome and predictors
outcome_var <- "COVID"
predictor_vars <- c("EXERCISE", "SLEEP", "DEPRESSED", "TROUBLE_SLEEPING", "SOCIAL")

# Loop over imputations and compute AUC
auc_list <- map(data_imputed, function(dat) {
  # Drop missing in predictors or outcome if any slipped through
  dat <- dat %>% drop_na(all_of(c(outcome_var, predictor_vars, "SAMPWEIGHT")))

@@

  # Fit random forest model (instead of lasso, because it's a classification problem)
  rf_model <- ranger(
    formula = as.formula(paste(outcome_var, "~", paste(predictor_vars, collapse = "+"))),
    data = dat[train_idx, ],
    case.weights = w_train,
    probability = TRUE
  )

  # Predict probabilities for class 1
  pred_probs <- predict(rf_model, data = dat[test_idx, ])$predictions[,2]

  # Compute AUC
  auc_result <- roc(y_test, pred_probs)$auc
  return(auc_result)
})

# Pool AUC across imputations
mean_auc <- mean(unlist(auc_list))

# Output
cat("AUC for each imputation:\n")
print(unlist(auc_list))
cat("\nPooled AUC across imputations:", round(mean_auc, 3), "\n")
```

Finally, we will construct a random forest model. For this, I will not use the multiple imputation dataset, as before, I will use the first imputed dataset, with cross validation 5 folds. The pooled AUC is 0.556.

```{r}
library(ranger)
library(caret)
library(pROC)
library(dplyr)

set.seed(1234)
k <- 5
auc_list <- numeric(length(data_imputed))

for (i in seq_along(data_imputed)) {
  data1 <- data_imputed[[i]]

  # Remove missing outcome
  data1_complete <- data1 %>% filter(!is.na(COVID))

  # Make sure outcome is binary (0/1)
  data1_complete$COVID <- as.numeric(data1_complete$COVID)

  # Add weights column if not present
  data1_complete$weights <- data1_complete$SAMPWEIGHT

  # Create folds
  folds <- createFolds(data1_complete$COVID, k = k, list = TRUE, returnTrain = FALSE)

  auc_folds <- numeric(k)

  for (fold_idx in seq_along(folds)) {
    test_idx <- folds[[fold_idx]]
    train_idx <- setdiff(seq_len(nrow(data1_complete)), test_idx)

    train_data <- data1_complete[train_idx, ]
    test_data  <- data1_complete[test_idx, ]

    # Fit random forest classifier
    rf_model <- ranger(
      formula = COVID ~ EXERCISE + SLEEP + DEPRESSED + TROUBLE_SLEEPING + SOCIAL, #formula
      data = train_data,
      case.weights = train_data$weights,
      probability = TRUE
    )

    # Predict probabilities for class 1
    pred_probs <- predict(rf_model, data = test_data)$predictions[,2]

    # Compute AUC
    auc_folds[fold_idx] <- roc(test_data$COVID, pred_probs, quiet = TRUE)$auc
  }

  # Store average AUC for this imputation
  auc_list[i] <- mean(auc_folds)
}

# Pool AUC across imputations
mean_auc <- mean(auc_list)
print(paste("Pooled AUC across imputations:", round(mean_auc, 3)))
```

Let's put these into a nice table for the manuscript. 
```{r}
# Create a data frame with the results
auc_table <- data.frame(
  Model = c("Multivariate", "Lasso", "Random Forest"),
  AUC = c(0.523, 0.529, 0.556)
)

# Create and style the GT table
model_table <- auc_table %>%
  gt() %>%
  tab_header(
    title = "Model Performance Comparison",
  ) %>%
  fmt_number(
    columns = vars(AUC),
    decimals = 3
  )

# Print the table
model_table

# Save the table as a PNG
here()
gtsave(model_table, here("results", "figures", "model_table.png"))
```
